Interisland movements of woodland caribou in Newfoundland

### 2020-02-19 paper/manuscript.Rmd
edit fogo info

Here, we describe the visual observation of an adult male caribou swimming
between islands in the Fogo Island archipleago in the North Atlantic Ocean. This
observation prompted us to query a large database of GPS radio-collared caribou
in the Fogo Island caribou herd in Newfoundland, Canada. We identified  86
remotely sensed swimming events between XX islands from three individual caribou
over a three-year period.

@@

Caribou were introduced to Fogo Island (n = 26 animals introduced between
1964–67) and Change Island (n = 5 animals introduced in 1964) from the main
Island of Newfoundland as part of a series of translocations and introductions
throughout the province (Bergerud & Mercer 1989). While caribou populations are
in decline throughout Canada (Festa-Bianchet et al. 2011), as well as in
Newfoundland (Bastille-Rousseau et al. 2013), the Fogo Island herd has not
declined to the same extent and currently consists of approximately 300 animals
(Newfoundland and Labrador Wildlife Division, unpublished data).

@@

We fitted 29 (n = 15 in 2016; n = 14 in 2017) adult female caribou on Fogo
Island (n = 25) and Western Indian Island (n = 4) with GPS collars (1240g, GPS
4400M; Lotek Wireless Inc., Newmarket, Ontario, Canada). Collars were deployed
on individual caribou for two years and collars were redeployed on the same
individuals for the duration of our study. Collars were programmed to collect
relocation fixes every 2 hours. All animal capture and handling procedures were
consistent with the American Society of Mammologists guidelines (Sikes & Gannon
2011) and were approved by Memorial University Animal Use Protocol No. 20152067.
During routine fieldwork on May 30, 2017, we observed an unmarked male caribou
swim between Perry and Eastern Indian Islands, a swim of at least 470m which
took ~9 minutes (~52m/minute). To determine the prevalence of ocean swimming for
the Fogo Island caribou herd, we queried nearly 200,000 GPS relocations from
26/29 collared adult female caribou between April 2016 and March 2019. We
determined all instances of individual caribou moving between islands. We
restricted the GPS dataset to the ice-free months of the year (beginning of
April to end of November) and we assumed caribou swam in the ocean between
islands when consecutive relocation fixes, i.e., 2 hour intervals, were located
on different islands. For each swimming event, we identified the origin island,
the terminal island, and the duration of time spent on the terminal island
before swimming again.

>>>>

Here, we describe the visual observation of an adult male caribou swimming
between islands in the Fogo Island archipleago in the North Atlantic Ocean.
During routine fieldwork on May 30, 2017, we observed an unmarked adult male
caribou swim between Perry and Eastern Indian Islands (Figure 1), a swim of at
least 470 m which took ~9 minutes (~52 m per minute). This observation prompted
us to query a large database of GPS radio-collared caribou in the Fogo Island
caribou herd in Newfoundland, Canada. We identified  86 remotely sensed swimming
events between XX islands from three individual caribou over a three-year
period.

Caribou were introduced to Fogo Island (n = 26 animals introduced between
1964–67) and Change Island (n = 5 animals introduced in 1964) from the main
Island of Newfoundland as part of a series of translocations and introductions
throughout the province (Bergerud & Mercer 1989). While caribou populations are
in decline throughout Canada (Festa-Bianchet et al. 2011), as well as in
Newfoundland (Bastille-Rousseau et al. 2013), the Fogo Island herd has not
declined to the same extent and currently consists of approximately 300 animals
(Newfoundland and Labrador Wildlife Division, unpublished data). Since 2016 we
have been studying the Fogo Island caribou herd (Bonar et al. 2017; Bonar et al.
2018; Peignier et al. 2019). To date, we have fitted XX adult female caribou
with GPS collars (1240g, GPS 4400M; Lotek Wireless Inc., Newmarket, Ontario,
Canada), which were programmed to collect relocation fixes every 2 hours (see
Bonar et al. 2018 for details).

@@

@@
